---
title: "CAT simulation"
author: "aiden"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Welcome

This page will give you a general idea of how to simulate a CAT using the `catR` package. 

The developers has made it really easy for us to run simulations. 

I am still building the content for this page. So please bear with it.

However, with the limited information you can still run the code and try it yourself. 

## Load catR package

```{r load package, message= FALSE}
require(catR)
```

## Run CAT simulation
```{r intro to CAT}
### Load 'polyIQ' item bank
bank <- read.table("https://www.dropbox.com/s/ml920oyi1z2npid/polyIQ.txt?raw=1",header = TRUE)
bank <- as.matrix(bank)
head(bank)

### CAT design:
#creation of logic for polytomous is exactly the same for dichtomous items.
## two starting items
startList <- list(nrItems = 2, theta=0)

## next item selection by 'MFI', EAP ad-interim proficiency estimation
testList <- list(method = "EAP", itemSelect = "MFI")

## stop after 10 items
stopList <- list(rule = "length", thr = 10)

## final proficiency estimation using maximum likelihood
finalList <- list(method = "BM")

### generation of a CAT response pattern
res <- randomCAT(trueTheta = 0, itemBank = bank, model="PCM",
                 start = startList, test = testList, stop = stopList, final = finalList)

## display of the output
res

```

## Plot Simulation
```{r Plot CAT, echo=FALSE}
## graphical display of the output
plot(res)
plot(res,ci=TRUE)

```





<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-98916204-1', 'auto');
  ga('send', 'pageview');

</script>