---
title: "Text Analysis"
author: "Aiden Loe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  #word_document:
    #toc: yes
    #toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Introduction**
Hello! I am listing down things that I found difficult when it comes to text analysis. 

## **White and Blank spaces**
This took me a while to learn. Hint** white spaces and blank spaces are not the same. 

`trimsw()` will not work at all if it is blank spaces. 


Here are some useful tips to know more about regular expressions.

+ [:digit:] or \d: digits, 0 1 2 3 4 5 6 7 8 9, equivalent to [0-9].

+ \D: non-digits, equivalent to [^0-9].

+ [:lower:]: lower-case letters, equivalent to [a-z].

+ [:upper:]: upper-case letters, equivalent to [A-Z].

+ [:alpha:]: alphabetic characters, equivalent to [[:lower:][:upper:]] or [A-z].

+ [:alnum:]: alphanumeric characters, equivalent to [[:alpha:][:digit:]] or [A-z0-9].

+ \w: word characters, equivalent to [[:alnum:]_] or [A-z0-9_].

+ \W: not word, equivalent to [^A-z0-9_].

+ \s[:xdigit:]: hexadecimal digits (base 16), 0 1 2 3 4 5 6 7 8 9 A B C D E F a b c d e f, equivalent to [0-9A-Fa-f].

+ "[:blank:]": blank characters, i.e. space and tab.

+ "[:space:]": space characters: tab, newline, vertical tab, form feed, carriage return, space.

+ \s: space, ` `.

+ \S: not space.

+ [:punct:]: punctuation characters, ! " # $ % & ’ ( ) * + , - . / : ; < = > ? @ [  ] ^ _ ` { | } ~.

+ [:graph:]: graphical (human readable) characters: equivalent to [[:alnum:][:punct:]].

+ [:print:]: printable characters, equivalent to [[:alnum:][:punct:]\\s].

+ [:cntrl:]: control characters, like \n or \r, [\x00-\x1F\x7F].


#### Some stuff worth remembering. 


If you want to find out if multiple strings exists.

```{r}
Text <- c('I went out to meet some friends')
words <- unlist(strsplit(Text, split = " "))
grep('to|meet', words, value=T) # actual matches
grepl('to|meet', words) # logical 

```

If you want to remove the rest of a string after a specific character, you can use the `gsub()` function

+ `.` matches any single character after the specific character 
+ `*` matches at least 0 times. 
+ `\` suppress the special meaning of metacharacters in regular expression. It has to be double backslash (i.e. `\\`) since `\` itself needs to be escaped in R. 

```{r}

text <- c('dd1 =~aaa', 'dd2 =~`bbb')
trimws(gsub("\\=~.*", "", text)) 

```

If you want to replace multiple patterns in a single string,

```{r}
text <- c('I am having a lot of fun')
gsub('having|fun','cool' ,text)
```


If you want to keep everything within the special characters only. It doesn't work if you have special characters before or after the opening and closing brackets though.

```{r}
item <- 'I like to [A1], [A2], [A3], [K1] fun'
words <- unlist(strsplit(item, split= ' '))
position <- (grep(c('\\['), words))
features_special_characters <- words[position]
(features <- gsub('\\[|\\]', "", features_special_characters))

```


#### Citations

+ http://stat545.com/block022_regular-expression.html



<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-98916204-1', 'auto');
  ga('send', 'pageview');

</script>