---
title: "Web Scrapping"
author: "Aiden Loe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Web Scrapping (Part 1)**

Scrapping viewership is one area of web scrapping, but perhaps you might be interested in doing sentiment analysis on content. 

So we want to extract the contents of the web pages rather than number of times someone viewed the web page. 

## **Scrape content (Wiki)**

We will be using the `RCurl` and `XML` package to help us with the scrapping. 

Let's use the Eurovision_Song_Contest as an example. 

The `XML` package has plenty functions that can allow us to scrape the data. 

Usually we are extracting information based on the tags of the web pages. 


```{r Scrape Content (wiki), message=FALSE, results='hide', warning=FALSE}
##### SCRAPPING CONTENT OFF WEBSITES ######
require(RCurl)
require(XML)
# XPath is a language for querying XML 
# //Select anywhere in the document
# /Select from root
# @select attributes. Used in [] brackets

#### Wikipedia Example ####
url <- "https://en.wikipedia.org/wiki/Eurovision_Song_Contest"
txt = getURL(url) # get the URL html code

# parsing html code into readable format
PARSED <- htmlParse(txt)

# Parsing code using tags
xpathSApply(PARSED, "//h1")

# strops code and return content of the tag
xpathSApply(PARSED, "//h1", xmlValue) # h1 tag
xpathSApply(PARSED, "//h3", xmlValue) # h3 tag
xpathSApply(PARSED, "//a[@href]") # a tag with href attribute


# Go to url 
# Highlight references
# right click, inspect element
# Search for tags
xpathSApply(PARSED, "//span[@class='reference-text']",xmlValue) # parse notes and citations
xpathSApply(PARSED, "//cite[@class='citation news']",xmlValue) # parse citation news
xpathSApply(PARSED, "//span[@class='mw-headline']",xmlValue) # parse headlines
xpathSApply(PARSED, "//p",xmlValue) # parsing contents in p tag
xpathSApply(PARSED, "//cite[@class='citation news']/a/@href") # parse links under citation. xmlValue not needed. 
xpathSApply(PARSED, "//p/a/@href") # parse href links under all p tags
xpathSApply(PARSED, "//p/a/@*") # parse all atributes under all p tags

# Partial matches - subtle variations within or between pages. 
xpathSApply(PARSED, "//cite[starts-with(@class, 'citation news')]",xmlValue) # parse citataion news that starts with..
xpathSApply(PARSED, "//cite[contains(@class, 'citation news')]",xmlValue) # parse citataion news that contains.

# Parsing tree like structure
parsed<-   htmlTreeParse(txt, asText = TRUE)

```

## **Scrape content (BBC)**

When you know the structure of the data. 

All you need to do is to find the correct function to scrape. 

```{r Scrape Content (BBC), message=FALSE, results='hide', warning=FALSE}


##### BBC Example ####
url <- "https://www.bbc.co.uk/news/uk-england-london-46387998"
url <- "https://www.bbc.co.uk/news/education-46382919"
txt = getURL(url) # get the URL html code

# parsing html code into readable format
PARSED <- htmlParse(txt)
xpathSApply(PARSED, "//h1", xmlValue) # h1 tag
xpathSApply(PARSED, "//p", xmlValue) # p tag
xpathSApply(PARSED, "//p[@class='story-body__introduction']", xmlValue) # p tag body
xpathSApply(PARSED, "//div[@class='date date--v2']",xmlValue) # date, only the first is enough
xpathSApply(PARSED, "//meta[@name='OriginalPublicationDate']/@content") # sometimes there is meta data. 
```

## **Create simple BBC scrapper**

Sometimes, creating a function will make your life better and make your script look simpler.  

```{r Plot BBC scrapper}
##### Create simple BBC scrapper #####
# scrape title, date and content
BBCscrapper1<- function(url){
  txt = getURL(url) # get the URL html code
  PARSED <- htmlParse(txt) # Parse code into readable format
  title <- xpathSApply(PARSED, "//h1", xmlValue) # h1 tag
  paragraph <- xpathSApply(PARSED, "//p", xmlValue) # p tag
  date <- xpathSApply(PARSED, "//div[@class='date date--v2']",xmlValue) # date, only the first is enough
  date <- date[1]
  return(cbind(title,date))
  #return(as.matrix(c(title,date)))
}


# Use function that was just created. 
BBCscrapper1("https://www.bbc.co.uk/news/education-46382919")
```

## **Keeping it neat**

Using the `plyr` package helps to arrange the data in an organised way. 
```{r Keeping it neat, message=FALSE}
## Putting the title and date into a dataframe
require(plyr)
#url
url<- c("https://www.bbc.co.uk/news/uk-england-london-46387998", "https://www.bbc.co.uk/news/education-46382919")
## ldply: For each element of a list, apply function then combine results into a data frame
#put into a dataframe
ldply(url,BBCscrapper1)

```



# **Web Scrapping (Part 2)**
This example below is taken from code kindly written by David stillwell. 

Some editing has been made to the original code. 

## **Scrape from Wiki tables**

You have learned how to scrape viewership on wikipedia and content on web pages. 

This section is about scrapping data tables online. 

```{r Scrape Table (Wiki), message=FALSE}
# Install the packages that you don't have first. 
library("RCurl") # Good package for getting things from URLs, including https
library("XML") # Has a good function for parsing HTML data
library("rvest") #another package that is good for web scraping. We use it in the Wikipedia example

#####################
### Get a table of data from Wikipedia
## all of this happens because of the read_html function in the rvest package
# First, grab the page source
us_states = read_html("https://en.wikipedia.org/wiki/List_of_U.S._states_and_territories_by_population") %>% # piping
  # then extract the first node with class of wikitable
  html_node(".wikitable") %>% 
  # then convert the HTML table into a data frame
  html_table()

```

## **Scrape from online tables**

If we can have two data tables that have at least one column with the same name, then we can merge them together. 

The main idea is to link the data together to run simple analysis.

In this case we can get data about [funding](http://apps.saferoutesinfo.org/legislation_funding/state_apportionment.cfm) given to various US states to support building infrastructure to improve students' ability to walk and bike to school. 

```{r Scrape Table (non-wiki)}

######################
url <- "http://apps.saferoutesinfo.org/legislation_funding/state_apportionment.cfm"
funding<-htmlParse(url) #get the data

# find the table on the page and read it into a list object
funding<- XML::readHTMLTable(funding,stringsAsFactors = FALSE)

funding.df <- do.call("rbind", funding) #flatten data
# Contain empty spaces previously.
colnames(funding.df)[1]<- c("State") # shorten colname to just State. 

# Match up the tables by State/Territory names
# so we have two data frames, x and y, and we're setting the columns we want to do the matching on by setting by.x and by.y
mydata = merge(us_states, funding.df, by.x="State, federal district, or territory", by.y="State")
# it looks pretty good, but note that we're down to 50 US States, because the others didn't match up by name
# e.g. "District of Columbia" in the us_states data, doesn't match "Dist. of Col." in the funding data

#Replace the total spend column name with a name that's easier to use.
colnames(mydata)[18] = "total_spend"

#  We need to remove commas so that R can treat it as a number.
mydata[,"Population estimate, July 1, 2017[4]"] = gsub(",", "", mydata[,"Population estimate, July 1, 2017[4]"]) 
mydata[,"Population estimate, July 1, 2017[4]"] = as.numeric(mydata[,"Population estimate, July 1, 2017[4]"]) #this converts it to a number data type

# Now we have to do the same thing with the funding totals, which are in a format like this: $17,309,568
mydata[,"total_spend"] = gsub(",", "", mydata[,"total_spend"]) #this removes all commas
mydata[,"total_spend"] = gsub("\\$", "", mydata[,"total_spend"]) #this removes all dollar signs. We have a \\ because the dollar sign is a special character.
mydata[,"total_spend"] = as.numeric(mydata[,"total_spend"]) #this converts it to a number data type

# Now we can do the plotting
options(scipen=9999) #stop it showing scientific notation
plot(mydata[,"Population estimate, July 1, 2017[4]"], mydata[,"total_spend"])

## What's does the correlation between state funding and state population look like?
cor(mydata[,"Population estimate, July 1, 2017[4]"], mydata[,"total_spend"]) # 0.9924265 - big correlation!

```

## **Plot funding data on map**

Perhaps it might be more interesting to see how the data is like on a map. 

We can utilise `map_data` function in the `ggplot` package to help us with that. 

Again, with a bit of data manipulation, we can merge the data table that contains the longitude and latitude information together with the funding data across different states.

```{r Plot on Map, message=FALSE, fig.width=18, fig.height=11}

require(ggplot2)
all_states <- map_data("state") # states
colnames(mydata)[1] <- "state" # rename to states
mydata$state <- tolower(mydata$state) #set all to lower case
Total <- merge(all_states, mydata, by.x="region", by.y = 'state') # merge data
# we have data for delaware but not lat, long data in the maps
i <- which(!unique(all_states$region) %in% mydata$state) 

# Plot data
ggplot() + 
  geom_polygon(data=Total, aes(x=long, y=lat, group = group, fill=Total$total_spend),colour="white") + 
  scale_fill_continuous(low = "thistle2", high = "darkred", guide="colorbar") + 
  theme_bw()  + 
  labs(fill = "Funding for School" ,title = "Funding for School between 2005 to 2012", x="", y="") + 
  scale_y_continuous(breaks=c()) +
  scale_x_continuous(breaks=c()) +
  theme(panel.border =  element_blank(),
        text = element_text(size=20))
```


# **Web Scrapping (Part 3)**
## **Scrape from discussion forums**

Coming soon..



<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-98916204-1', 'auto');
  ga('send', 'pageview');

</script>
